---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nlmixr2save

<!-- badges: start -->
[![R-CMD-check](https://github.com/nlmixr2/nlmixr2save/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/nlmixr2/nlmixr2save/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/nlmixr2/nlmixr2save/graph/badge.svg)](https://app.codecov.io/gh/nlmixr2/nlmixr2save)
<!-- badges: end -->

The goal of nlmixr2save is to save a nlmixr2 fit in file formats that
can be read outside R, avoiding `.rds` or other formats as much as
possible.

## Installation

You can install the development version of nlmixr2save from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("nlmixr2/nlmixr2save")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}

library(nlmixr2)
library(nlmixr2save) # eventually nlmixr2 will include this if available
one.cmt <- function() {
  ini({
    ## You may label each parameter with a comment
    tka <- 0.45 # Log Ka
    tcl <- log(c(0, 2.7, 100)) # Log Cl
    ## This works with interactive models
    ## You may also label the preceding line with label("label text")
    tv <- 3.45; label("log V")
    ## the label("Label name") works with all models
    eta.ka ~ 0.6
    eta.cl ~ 0.3
    eta.v ~ 0.1
    add.sd <- 0.7
  })
  model({
    ka <- exp(tka + eta.ka)
    cl <- exp(tcl + eta.cl)
    v <- exp(tv + eta.v)
    linCmt() ~ add(add.sd)
  })
}

fitF <- suppressMessages(nlmixr(one.cmt, theo_sd, est="focei",
                                control=list(print=0, compress=FALSE)))

saveFit(fitF)

# You can also load this with `loadFit(fitF)`

print(loadFit(fitF))
```

This creates a ".zip" file that includes a "fitF.R" file.  This fit
builds the fit from (mostly) csv and R source files.

The reason for this extra package saving models:

- The model properties and results can be examined outside of R

- This model can be saved without worrying how data are serialized and/or compressed

- This saved model version **should** be independent of the version of
  `rxode2` and `nlmixr2`, whereas other methods of saving are
  dependent on the version of nlmixr2/rxode2

In the future, this could be used in conjunction with `vroom`, `altrep` and other
similar tools to partially load fit without loading the corresponding datasets.

In addition to introducing the R-independent zip file for nlmixr2
fits, this also introduced the cache assignment operator `:=` which
will cache functions calls when assigning them.

In the case of `nlmixr2` fits, these cached values will be the zip
files described above; Otherwise, this will be a `rds` file that has
the object and the md5 of the arguments in a list.

If the object is NOT a function, it cannot load the cache, but may
have some side effects.

In the case of a nlmixr2 fit, you could assign it as follows

``` r
fit2 := fit # The fit2.zip is created to save the fit
```

Here are some annotated examples of the cache assignment operator in
pratice:

```{r example2}
# You can also automatically load a fit if the arguments for nlmixr2 do not change
# by using the := cached assignment operator introduced in this package

fitF := nlmixr(one.cmt, theo_sd, est="focei", control=list(print=0, compress=FALSE))

# Now using the same arguments it loads
fitF := nlmixr(one.cmt, theo_sd, est="focei", control=list(print=0, compress=FALSE))

# But if we use different arguments, say using saem, it will rerun.
fitF := nlmixr(one.cmt, theo_sd, est="saem", control=list(print=0, compress=FALSE))

# note this approach does not evaluate the first function only,
# everything else is evaluated so `suppressMessage(nlmixr(...))` would always run
# the model, never loading the cached model.

# The cached assignment operator works for generic functions too, but saves
# as a rds file.  For example,

sim := rxSolve(one.cmt, theo_sd) # First simulate

sim := rxSolve(one.cmt, theo_sd) # Same arguments, load instead

sim := rxSolve(one.cmt, theo_sd, nStud=4) # Different arguments, resimulate

```
