---
title: "Using nlmixr2save"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using nlmixr2save}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`nlmixr2save` focuses on two related problems:

1. Saving `nlmixr2` results in a format that stays readable outside of
   `.rds`.

2. Reusing expensive fits or simulations when nothing important has changed.

## Why save a fit this way?

`saveFit()` writes a saved fit as a collection of files and then optionally zips
them together.  In practice, most of the saved object is reconstructed from:

- `.R` files for model definitions and objects that can be recreated
  as source.

- `.csv` files for tabular fit components and datasets.

That means the saved fit is largely inspectable outside of R, and it is not tied
to the binary serialization format used by a specific version of `nlmixr2` or
`rxode2`.

```{r, eval = FALSE}
library(nlmixr2est)
library(nlmixr2data)
library(nlmixr2save)

fit <- nlmixr2(one.cmt, theo_sd, est = "focei")
saveFit(fit, "fit")

restored_fit <- loadFit("fit")
```

When `loadFit("fit")` runs, it recreates the object by sourcing the
generated `.R` (for fit this would be `fit.R`) files and reading the
generated `.csv` files back in.

This is the main protection against a saved fit becoming unreadable
simply because an internal serialization format changes.

## What gets restored?

For deterministic estimation methods, the restored object includes the full
saved fit, including the original model, fit results, and `origData`.

This is especially useful for long-running estimation jobs because the saved fit
can be rebuilt without repeating the estimation itself.

## Dataset-aware caching

This `saveFit()` can automatically be performed and cached with a new
operator `:=` and is the quickest way to save nlmixr2 fits (and other
items).  For example:

```{r, eval = FALSE}
fit := nlmixr2(one.cmt, theo_sd, est = "focei")
```

For `nlmixr2` fits, the cache key is based on:

- the normalized model definition,
- the estimation method,
- the control and table options, and
- a simplified version of the dataset that keeps the standard estimation
  columns, covariates, and requested `table$keep` columns.

This has two important consequences.

### 1. Irrelevant dataset changes do not force a refit

If the new dataset only changes columns that are not used for estimation,
`nlmixr2save` restores the existing fit instead of rerunning the estimation.

```{r, eval = FALSE}
fit := nlmixr2(one.cmt, theo_sd, est = "focei")

theo_sd_extra <- theo_sd
theo_sd_extra$.ignored <- "notes"

fit := nlmixr2(one.cmt, theo_sd_extra, est = "focei")
```

The expensive estimation is skipped, but the restored object still updates
`fit$origData` to the new dataset.  In other words, the cached fit is reused
only when the meaningful estimation inputs match, while the saved object still
tracks the latest original dataset you supplied.

### 2. Meaningful dataset changes do force a refit

If you change a column that matters to the fit, such as `DV`, time, dosing
information, a covariate, or a `table$keep` column, the cache key changes and
the estimation is run again.

This is the intended safety boundary: harmless dataset changes are absorbed, but
real estimation changes invalidate the cache.

## Using `:=` for long-running fits and simulations

The `:=` operator caches the result under the object name on the left-hand side.
If the saved result matches the current call, it restores the cached object
instead of rerunning the call.

```{r, eval = FALSE}
fit := nlmixr2(one.cmt, theo_sd, est = "focei") # creates fit.zip

# Same call: restore from cache
fit := nlmixr2(one.cmt, theo_sd, est = "focei")

# Different estimation method: rerun
fit := nlmixr2(one.cmt, theo_sd, est = "saem") # overwrites fit.zip
```

For deterministic `nlmixr2` fits, the cached form is the text-and-csv-based fit
bundle described above.  For other functions, the cached form is usually an
`.rds` file.

## Seed-aware restores for stochastic work

Some calculations depend on the random-number stream.  For those, `:=` stores
both the result and random-state metadata.  When the result is restored, the
seed is advanced to the same post-run state so downstream code sees the same
random stream it would have seen if the expensive call had actually run.

This is what makes `:=` useful for long-running simulations and stochastic
estimation methods: you can restore the result without silently changing the
reproducibility of the rest of the script.

## Integrating `nlmixr2save` into your package

There are two common integration paths.

### Deterministic estimation

If your package returns standard `nlmixr2` fit objects through a deterministic
estimation method, users can already write:

```{r, eval = FALSE}
fit := nlmixr2(model, data, est = "focei") # saves fit
```

and get zip-based save/restore behavior automatically.

### Stochastic estimation or simulation

If your package provides a stochastic workflow, use the seed-aware path instead.
For plain simulation functions, register the function name with
`saveFitRandom()`.  For `nlmixr2` estimation methods, mark the estimator as
random so `:=` knows to use the seed-aware cache path.

The companion vignette
`vignette("register-simulation-functions", package = "nlmixr2save")` shows both
patterns.

## Limitations

`nlmixr2save` is intentionally conservative, and a few limitations are worth
keeping in mind:

1. The saved fit is primarily `.R` and `.csv`, but not exclusively.  Some fit
   components still fall back to `.rds` when they cannot be safely recreated as
   text.

2. Cache reuse only works when `:=` sees the expensive call directly.  Wrapping
   the call inside something like `suppressMessages(nlmixr2(...))` forces the
   call to run before caching can intercept it.

3. If dataset simplification cannot be computed, caching falls back to hashing
   the full dataset.  That is safe, but it can cause more reruns than strictly
   necessary.

4. Seed-aware restores require the same starting random state.  If the seed is
   different, the cached stochastic result is discarded and rerun.

5. The cache files are named from the left-hand-side object name and written in
   the current working directory, so project-level file management still
   matters.
