---
title: "Tracking the nlmixr2est and rxode2 versions of a saved fit"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tracking the nlmixr2est and rxode2 versions of a saved fit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

A saved `nlmixr2` fit is the result of two packages working together:
`nlmixr2est` estimates the model, and `rxode2` compiles and solves it.  A change
in either package can, in principle, change the fit.  Because a saved fit is
meant to be reloaded later -- possibly months later, on a different machine, with
newer packages installed -- `nlmixr2save` records which versions produced it and
tells you when they no longer match.

This behavior is easy to miss, because when the versions *do* match (the common
case) nothing is printed at all.

## What is recorded

When you `saveFit()` a fit (or cache one with `:=`), `nlmixr2save` stores a small
piece of metadata alongside it: the `nlmixr2est` and `rxode2` **versions**, and,
when either package was installed from a remote such as GitHub (via `remotes` or
`pak`), its **commit sha**.  The metadata is embedded in the saved fit's loader
script, so it travels inside the portable `.zip` and survives a load -> save
round-trip (it always records the version that *ran* the fit, not the version
that happened to re-save it).

Nothing about your model or data is stored here -- only the package versions.

## What happens on load

### `loadFit()`

When you load a fit and the installed `nlmixr2est` or `rxode2` differs from the
version that produced it, `loadFit()` warns:

```{r, eval = FALSE}
fit2 <- loadFit("fit")
#> Warning: this fit was run with nlmixr2est 6.2.0 (installed 6.3.0)
```

The fit still loads normally -- the warning is only there to explain why a
reloaded fit might not be bit-for-bit identical to one you would get by
re-running today.  `loadFit()` cannot re-run the fit itself, because a saved fit
does not carry the original `nlmixr2()` call.

### The `:=` caching operator

The `:=` operator *does* have the original call, so it can offer to re-run.  When
a cached fit was produced by a different package version:

- In an **interactive** session, you are asked what to do:

  ```{r, eval = FALSE}
  fit := nlmixr2(one.cmt, theo_sd, est = "focei")
  #> The cached fit was run with nlmixr2est 6.2.0 (installed 6.3.0).
  #>
  #> 1: Reload the cached fit as-is
  #> 2: Rerun the fit with the installed packages
  ```

  Choosing **2** re-runs the fit with the currently installed packages and
  refreshes the cache; choosing **1** keeps the cached fit.

- In a **non-interactive** session (scripts, `R CMD build`, CI, rendering a
  document), there is no one to ask, so `:=` loads the cached fit and warns:

  ```
  #> Warning: the cached fit was run with nlmixr2est 6.2.0 (installed 6.3.0);
  #> loading the cached fit
  ```

If both packages changed, the message names both, e.g.
`nlmixr2est 6.2.0 (installed 6.3.0); rxode2 5.1.3 (installed 6.0.0)`.

## Turning the check on or off

The version check is controlled by a single option, `nlmixr2save.checkVersion`,
which defaults to `TRUE`:

```{r, eval = FALSE}
# Silence version-mismatch warnings and skip the interactive rerun prompt
options(nlmixr2save.checkVersion = FALSE)

# Turn the check back on (the default)
options(nlmixr2save.checkVersion = TRUE)
```

`loadFit()` also takes a per-call `checkVersion` argument, which defaults to the
option, so you can override it for a single load without changing the global
setting:

```{r, eval = FALSE}
fit2 <- loadFit("fit", checkVersion = FALSE) # load this one quietly
```

A common place to set the option is at the top of a script or vignette that
deliberately ships a committed cache and does not want version warnings:

```{r, eval = FALSE}
options(nlmixr2save.checkVersion = FALSE)
```

### Relationship to trusted-cache mode

If you already run with `options(nlmixr2save.check = FALSE)` (trusted-cache
mode, described in `vignette("uses")`), `:=` loads a committed cache without any
model/data/version comparison at all.  Trusted-cache mode is intended to keep a
committed cache stable across package versions, so it never prompts or warns
about versions -- it is a superset of `nlmixr2save.checkVersion = FALSE` for the
`:=` path.  The `nlmixr2save.checkVersion` option is what you want when you are
using ordinary (checked) caching or `loadFit()` and simply want to control the
version comparison on its own.

## Backward compatibility

Fits saved by versions of `nlmixr2save` that predate this feature carry no
version metadata.  Loading them is completely silent: with nothing to compare
against, `nlmixr2save` never warns and never prompts.
